---
title: PML report
author: "by Vishnu"
output:
  html_document:
    fig_height: 9
    fig_width: 9
---


## Preprocessing of Data 
```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
 
### Data Reading
we can read the two csv files into two data frames.
```{r, cache = T}
trainRaw <- read.csv("./data/p-training.csv")
testRaw <- read.csv("./data/p-testing.csv")
dim(trainRaw)

dim(testRaw)
``` 

### Data Cleaning
Claening gets rid of missing valus of some observations and meaning less values.
```{r, cache = T}
sum(complete.cases(trainRaw))
```
Removing observations with no values
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  
getting rid of pointless values
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

### Data Slicing	
 Splitting data into standard 70 and 30.
```{r, cache = T}
set.seed(22328)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Modeling of Data
predictive model for activity recognition using Random forest algorithm with 5fold cross validation.	
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
estimating performance of model on validation sets.
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```
estimated accuracy of the model = 99.43% 
the estimated out-of-sample error = 0.58%.

## Test data prediction 
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Figures
1.Visualization of correalation matrix
```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```
2.Visualization of decision tree
```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```
